---
title: "assignment 3 - Draft 1"
author: "Claire Powers"
date: "5/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Santa Barbara County wants to reduce the risk of whale strikes by 60%

- Cost of implementation is $7 million

```{r message=FALSE, warning=FALSE}
library(tidyverse)
data <- read_csv("HW3Data-Numeric.csv")
```

####Problem 1. Show the model and interpret regression coefficients
```{r}
summary(data)

data$risk <- as.numeric(data$risk)
levels(data$risk)

data$bid <- as.numeric(data$bid)
levels(data$bid)

data$NEP <- as.numeric(data$NEP)
summary(data$NEP)

data$income <- as.numeric(data$income)
#levels(data$income) <- c("poor","middle","rich","very_rich","one_percent")
levels(data$income)

data$age <- as.numeric(data$age)
levels(data$age)

#data$vote <- as.factor(data$vote)

summary(data)

#Actual income

whale_lm <- lm(vote ~ risk + bid + NEP + income + age, data = data)
summary(whale_lm)
```

####1b. Interpreting coefficients
```{r}
#Regression coefficients - 
# Risk: Positive relationship
# Bid: Negetive relationship
# NEP: Positive relationship
# Income: Positive relationship
# Age: Negative relationship
```

####Value of a single prevented whale death
```{r}
summary(whale_lm)

Bid <- (.5 - (0.129 + .00076758*risk + 0.0158613*NEP + 0.0034322*income - 0.0035354*age))/(-.0010409)

bid20 <- (.5 - (0.129 + 0.00076758*20)/(-0.0010409))
bid20

bid24 <- (.5 - (0.129 + 0.00076758*24)/(-0.0010409))
bid24

bid24-bid20
```

```{r}
##New data with all bids = $60

risk60 <- as.numeric(rep(60,500))

data1 <- data.frame(risk60,data$bid,data$vote,data$NEP,data$income,data$age)
colnames(data1) <- c("risk","bid","vote","NEP","income","age")
```


####3. arbitrary respondents
```{r}
a <- sample_n(data1,1)
b <- sample_n(data1,1)
c <- sample_n(data1,1)

#Probability of a, b and c voting yes.
pred_a <- predict(whale_lm, newdata=a)
pred_b <- predict(whale_lm, newdata=b)
pred_c <- predict(whale_lm, newdata=c)

###a,b,c

WTPfun <- function(risk,NEP,income,age){
 bid <- (.5 - (0.129 + .00076758*risk + 0.0158613*NEP + 0.0034322*income - 0.0035354*age))/(-.0010409)
 return(bid)
}

WTPfun(a$risk,a$NEP,a$income,a$age)
a
WTPfun(b$risk,b$NEP,b$income,b$age)
b
WTPfun(c$risk,c$NEP,c$income,c$age)
c
```

```{r}
WTP <- WTPfun(data1$risk,data1$NEP,data1$income,data1$age)
colnames(WTP) <- c("WTP")

summary(WTP)

meanWTP <- mean(WTP)
meanWTP
```

```{r}

Yintfun <- function(risk,NEP,income,age){
 bid <- (0 - (0.129 + .00076758*risk + 0.0158613*NEP + 0.0034322*income - 0.0035354*age))/(-.0010409)
 return(bid)
}

Yint <- Yintfun(data1$risk,data1$NEP,data1$income,data1$age)
Yint

meanYint <- mean(Yint)

.5*(.5)*(meanYint-meanWTP)

total_benefit <- 120.0884*150000
total_benefit

total_cost <- meanWTP*150000
total_cost
```


```{r probabilities for whole data frame}
#probability of voting yes from original data frame
probs <- data.frame(predict(blm, newdata = data, type="response"))

#Adding column with 60 risk only
data$risk2 <- as.factor(rep("60",length.out = 500))

data2 <- data %>% 
  select("risk2","bid","vote","NEP","income","age")

colnames(data2) <- c("risk","bid","vote","NEP","income","age")

#Probability of yes votes with 60% reduction
riskreduced <- data.frame(predict(blm, newdata = data2, type="response"))

probabilities <- data.frame(probs,riskreduced)
colnames(probabilities) <- c("original","risk60")

```

- 20% reduction saves 5 lives

```{r an interesting function}
probs <- data.frame(logit2prob(blm$coefficients))
```

#60% risk reduction for a,b,c
#a
#a_2 <- data.frame(as.factor("60"),as.factor("40"),as.factor("1"),as.numeric("41"),as.factor("very_rich"),as.factor("40"))
#colnames(a_2) <- c("risk","bid","vote","NEP","income","age") 

#b
#b_2 <- data.frame(as.factor("60"),as.factor("100"),as.factor("0"),as.numeric("31"),as.factor("middle"),as.factor("30"))
#colnames(b_2) <- c("risk","bid","vote","NEP","income","age") 

#c
#c_2 <- data.frame(as.factor("60"),as.factor("20"),as.factor("1"),as.numeric("41"),as.factor("middle"),as.factor("50"))
#colnames(c_2) <- c("risk","bid","vote","NEP","income","age") 

#pred_a_2 <- predict(blm, newdata=a_2, type="response")
#pred_b_2 <- predict(blm, newdata=b_2, type="response")
#pred_c_2 <- predict(blm, newdata=c_2, type="response")

